### 2015-07-14 zplus.tex
Limitiation to three classes --- text in
Typo in table
Address question about only three levels of posterior surprise.

Under \zplus\ we may apply the normal operation
of probability theory
%% TODO Restore for QR'15
% --- conditionalization, Bayes' law, etc. ---
but the arithmetic operations we use for them change.  Rather than multiplying
probabilities, we add degrees of surprise%
%% TODO Restore for QR'15, but it's in the table anyway
%; rather than adding
%probabilities, we take the minimum surprise%
.  Goldszmidt and
Pearl~\shortcite{Goldszmidt:96} provide the following substitutions
between the quantitative probability function $P$ and its qualitative
counterpart $\kappa$ for formulas $\phi$ and $\psi$, and primitive events, $e$.
\begin{displaymath}
  \small
  \begin{array}{|l|l|}
    \hline
  P(\psi) = \sum_{e \in \psi} P(e) & \kappa(\psi) =
  \operatorname{min}_{e \in \psi} \kappa(\phi) \\ \hline
  P(\psi) + P(\neg \psi) = 1 & \kappa(\psi) = 0 \vee \kappa(\neg \psi) = 0
  \\ \hline
  P(\psi | \phi) = P(\psi \wedge \phi) / P(\phi) &
  \kappa(\psi | \phi) = \kappa(\psi \wedge \phi) - \kappa(\phi)
  \\ \hline
\end{array}
\end{displaymath}
\noindent
In a Bayes network, with conditionally independent $\omega$ and
$\phi$ we further have $\kappa(\omega \wedge \phi) = \kappa(\omega) + \kappa(\phi)$.
\hide{Instead of the probability
of an event being the sum of the probabilities of the primitive outcomes that
make up that event, the degree of surprise of an event is the minimum of the
degrees of surprise of the primitive outcomes that make it up. Instead of having
the probabilities of mutually exclusive and exhaustive events sum to one, at
least one of a set of mutually exclusive and exhaustive events must be
unsurprising. Finally, we have an analog of Bayes' law in which the normalizing
operation consists of subtraction rather than division.
}

@@

MIFD's assessment ranks hypotheses as either
\emph{likely}, \emph{plausible} or \emph{unlikely}. A hypothesis $h$,
conditioned on evidence $\omega$,
is \emph{likely} if
$\kappa(h|\omega) < \kappa(\neg h|\omega)$, \emph{plausible} if $\kappa(h|\omega) = \kappa(\neg h|\omega)$
and \emph{unlikely} otherwise.

>>>>

Under \zplus\ we may apply the normal operation
of probability theory
%% TODO Restore for QR'15
% --- conditionalization, Bayes' law, etc. ---
but the arithmetic operations we use for them change.  Rather than multiplying
probabilities, we add degrees of surprise%
%% TODO Restore for QR'15, but it's in the table anyway
%; rather than adding
%probabilities, we take the minimum surprise%
.  Goldszmidt and
Pearl~\shortcite{Goldszmidt:96} provide the following substitutions
between the quantitative probability function $P$ and its qualitative
counterpart $\kappa$ for formulas $\phi$ and $\psi$, and primitive events, $e$.
\begin{displaymath}
  \small
  \begin{array}{|l|l|}
    \hline
  P(\psi) = \sum_{e \in \psi} P(e) & \kappa(\psi) =
  \operatorname{min}_{e \in \psi} \kappa(e) \\ \hline
  P(\psi) + P(\neg \psi) = 1 & \kappa(\psi) = 0 \vee \kappa(\neg \psi) = 0
  \\ \hline
  P(\psi | \phi) = P(\psi \wedge \phi) / P(\phi) &
  \kappa(\psi | \phi) = \kappa(\psi \wedge \phi) - \kappa(\phi)
  \\ \hline
\end{array}
\end{displaymath}
\noindent
In a Bayes network, with conditionally independent $\omega$ and
$\phi$ we further have $\kappa(\omega \wedge \phi) = \kappa(\omega) + \kappa(\phi)$.
\hide{Instead of the probability
of an event being the sum of the probabilities of the primitive outcomes that
make up that event, the degree of surprise of an event is the minimum of the
degrees of surprise of the primitive outcomes that make it up. Instead of having
the probabilities of mutually exclusive and exhaustive events sum to one, at
least one of a set of mutually exclusive and exhaustive events must be
unsurprising. Finally, we have an analog of Bayes' law in which the normalizing
operation consists of subtraction rather than division.
}

@@

% Extended this to address [[file:todo.txt::*Review%203:%20Limitation%20to%20three%20classes][Review 3: Limitation to three classes]]
\mifd's assessment ranks hypotheses as either
\emph{likely}, \emph{plausible} or \emph{unlikely}. A hypothesis $h$,
conditioned on evidence $\omega$,
is \emph{likely} if
$\kappa(h|\omega) < \kappa(\neg h|\omega)$, \emph{plausible} if $\kappa(h|\omega) = \kappa(\neg h|\omega)$
and \emph{unlikely} otherwise.
That is, a hypothesis is likely (unlikely) if some maximally likely scenario labels it as
true (false) , and no maximally likely scenario labels it as false (true).  A
hypothesis is plausible, in any other situation.
Note that this restriction to three classes applies to \mifd's output,
not its inputs --- the \tkappa\ values of \mifd's underlying model
range over arbitrary natural numbers as needed to distinguish
qualitatively distinct likelihoods.
It would be possible to alter the assessment process to provide more classes of
likelihood, using the difference in kappa ranks between the maximally likely
scenario in which a hypothesis is labeled as true (false) and the one in which
it is labeled as false (true).  We have not seen the need to do so and, indeed,
worry that additional precision would come at the cost of validity.
